### 2022-01-03 input/proposal.tex
wip

\begin{proof}
Stale control information (see Definition~\ref{def:lwo}) may impair
the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. We simplify the echo since $\alpha_c^3$ is already transiting
towards $a$. Finally, Figure~\ref{fig:problemD} shows that eventually,
$c$ stays in the deleted partition $P_a$ for not having received
$\delta_a$ that $b$ blocked earlier. The system does not reach
consistent partitioning. \end{proof}

As a cascading effect in Figure~\ref{fig:problemD}, not only $c$
wrongfully believes that it belongs to $a$'s partition when $a$
already deleted it, but it could contribute to its inconsistency by
not forwarding farther but up-to-date messages from actual sources.

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning. For instance, $c$ detects the
blocking of $\delta_a$ because it acknowledges that its parent $b$
delivered a stale notification $\alpha_c^1$, \ie between the delivery
of $\alpha_c^1$ and the delivery of $\delta_c$, $b$ blocked any other
$\delta$ and therefore $\delta_a$.  \TODO{Must take action because its
  parent has forwarded something already stale so previous cannot be
  good.}

@@

\begin{lemma}[\label{lem:detector}$\textup{FB}^+ \wedge \textup{FS}
  \implies$ Detector existence] Assuming
  $\textup{FB}^+$ and $\textup{FS}$, a chain of delivery on $\pi_{xz}$ of
  $\alpha_x$ with $\alpha_x = \last D_z$, if $x$ broadcasts a staleness
  notification $\delta_x$, either $z$ eventually delivers it, or a
  \process $y$ in such a chain of delivery $\pi_{xz}'$ eventually
  acknowledges its blocking.
  % $b_p(\delta_x) \implies d_q(x) U Detect[X]$ ??? 
\end{lemma}

\begin{proof}
  %% knowing that b_p delta is sufficient to say that processes will
  %% receive delta or detect blocking
  (A) Assuming a chain of delivery $\pi_{xz} = [x, \ldots, z]$ of
  $\alpha_x$ with $\forall y \in \pi_{xz}: \alpha_x = \last D_y$, and
  assuming $d_x(\delta_x)$, then with forwarding of staleness
  (Definition~\ref{def:fs}),
  $f_{\pi_{xz}[k]}(\delta_x) \implies \eventually
  d_{\pi_{xz}[k+1]}(\delta_x)$ except if $\delta_x \in D_{\pi_{xz}[k+1]}$ which
  means that $\pi_{xz}[k+1]$ already delivered and forwarded $\delta_x$
  following another such a chain of delivery. Therefore, $z$
  eventually receives, hence delivers $\delta_x$.

  \noindent (B) Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \last D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $ y = \pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the forwarding stops when $y$ already
  delivered $\delta_x$ ($\delta_x \in D_y$ covered by (A)) or
  delivered a better message $\alpha_s = \last D_y$ with
  $\alpha_s \neq \alpha_x$. With $y'=\pi_{xz}[k+1]$, forwarding of
  best (Definition~\ref{def:fbplus}) implies three possible outcomes:
  \begin{enumerate}[(i)]
  \item $\alpha_{s'} = \last D_{y'}$ with $\alpha_{s'} \neq \alpha_x$
    hence $y' \iff y$ which leaves two possible outcomes:
  \item
    $r_{y'}(\alpha_s) \wedge \alpha_x = \last D_{y'} \wedge \alpha_x <
    \alpha_s$ which means that a shorter path of delivery $\pi_{xz}'$
    exists that either forwards appropriate staleness notifications
    (A) or detects its possible blocking:
  \item
    $r_{y'}(\alpha_s) \wedge \delta_s \in D_{y'} \wedge \alpha_x =
    \last D_{y'}$ which detects a possible blocking of
    $\delta_x$. Without global knowledge, $y'$ assumes it belongs to
    the shortest and only path of delivery of $\alpha_x$
    ($\pi_{xy'} \subseteq \Pi_{xz}$) thus it does not further delegate
    the detection to another \process.
  \end{enumerate} \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}

>>>>

\begin{proof}
Stale control information (see Definition~\ref{def:lwo}) may impair
the propagation of both
\begin{inparaenum}[(i)]
\item notifications about actual sources, and
\item notifications about deleted partitions.
\end{inparaenum}
Figure~\ref{fig:problem} depicts a scenario comprising only three
\processes $a$, $b$, $c$ chained with FIFO links, \ie where \processes
receive the messages in the order of their sending ($s_{pq}(m)
\rightarrow s_{pq}(m') \implies r_q(m) \rightarrow r_q(m')$). In
Figure~\ref{fig:problemA}, both $a$ and $c$ become sources, sending
their respective notification message to $b$. In
Figure~\ref{fig:problemB}, both $a$ and $c$ delete their partition
while $b$ receives, delivers, and forwards $\alpha_a^2$. In
Figure~\ref{fig:problemC}, $b$ receives, delivers, and forwards
$\alpha_c^1$, for it improves its best partition. Then it receives and
discards $\delta_a$, for its best partition does not match the deleted
one. It echoes back $\alpha_c^3$ to $a$. In Figures~\ref{fig:problemD}
and \ref{fig:problemE}, $a$ and $b$ handle transiting
notifications. Finally, Figure~\ref{fig:problemF} shows that
eventually, $c$ stays in the deleted partition $P_a$ for not having
received $\delta_a$ that $b$ blocked earlier. The system does not
reach consistent partitioning. \end{proof}

The issue is that each \process trusts its parent to forward all
staleness notifications relevant to it. When this fails, as in
Figure~\ref{fig:problem}, not only a \process ($c$) may wrongfully
stay in a partition ($P_a$) when its source ($a$) already deleted it,
but it may contribute to its inconsistency by not forwarding farther
but up-to-date messages from live sources.

A first step to avoid staying in inconsistent state is to extend the
behavior of each \process so \processes such as $c$, that would remain
in a wrong partition, use their history of receipt and delivery to
detect the blocking of staleness notifications that can hinder
reaching consistent partitioning.  In Figure~\ref{fig:problem}, $b$
blocked the staleness notification $\delta_a$ that $c$ needs since it
belongs to a ephemeral partition $P_c$ that $c$ acknowledges to be
stale before $b$ does. Therefore, $c$ can detect the blocking of
$\delta_a$ since it acknowledges that its parent $b$ received,
delivered, and forwarded the stale notification $\alpha_c^1$. In other
terms, $c$ acknowledges that between the delivery of $\alpha_c^1$ and
the delivery of $\delta_c$, $b$ blocked any other $\delta$ and
therefore, it could have blocked $\delta_a$.

@@

\begin{lemma}[\label{lem:detector}$\textup{FB}^+ \wedge \textup{FS}
  \implies$ Detector existence] Assuming
  $\textup{FB}^+$ and $\textup{FS}$, a chain of delivery on $\pi_{xz}$ of
  $\alpha_x$ with $\alpha_x = \last D_z$, if $x$ broadcasts a staleness
  notification $\delta_x$, either
  \begin{inparaenum}[(A)]
  \item\label{lem:detectorA}  $z$ eventually delivers it, or 
  \item\label{lem:detectorB} a \process $y$ in such a chain of
    delivery $\pi_{xz}'$ eventually acknowledges its blocking.
  \end{inparaenum}
  % $b_p(\delta_x) \implies d_q(x) U Detect[X]$ ??? 
\end{lemma}

\begin{proof}
  %% knowing that b_p delta is sufficient to say that processes will
  %% receive delta or detect blocking
  Assuming a chain of delivery $\pi_{xz} = [x, \ldots, z]$ of
  $\alpha_x$, we must prove that whatever the possible states of
  \nodes that belong to this chain, \TODO{WIP}
  \begin{asparadesc}
  \item [In the same partition:] $\forall y \in \pi_{xz}: \alpha_x =
    \last D_y$, and assuming $d_x(\delta_x)$, then with forwarding of
    staleness (Definition~\ref{def:fs}), $f_{\pi_{xz}[k]}(\delta_x)
    \implies \eventually d_{\pi_{xz}[k+1]}(\delta_x)$ except if
    $\delta_x \in D_{\pi_{xz}[k+1]}$ which means that $\pi_{xz}[k+1]$
    already delivered and forwarded $\delta_x$ following another such
    a chain of delivery. Therefore, $z$ eventually receives, hence
    delivers $\delta_x$.

  \item[ (\ref{lem:detectorB})] Assuming a chain of delivery
  $\pi_{xz} = [x, \ldots, z]$ of $\alpha_x$ with at least
  $\alpha_x = \last D_z$, and assuming $d_x(\delta_x)$, then by (A),
  the staleness notification reaches a \process $ y = \pi_{xz}[k]$. By
  Definition~\ref{def:fs}, the forwarding stops when $y$ already
  delivered $\delta_x$ ($\delta_x \in D_y$ covered by (A)) or
  delivered a better message $\alpha_s = \last D_y$ with
  $\alpha_s \neq \alpha_x$. With $y'=\pi_{xz}[k+1]$, forwarding of
  best (Definition~\ref{def:fbplus}) implies three possible outcomes:
  \begin{enumerate}[(i)]
  \item $\alpha_{s'} = \last D_{y'}$ with $\alpha_{s'} \neq \alpha_x$
    hence $y' \iff y$ which leaves two possible outcomes:
  \item
    $r_{y'}(\alpha_s) \wedge \alpha_x = \last D_{y'} \wedge \alpha_x <
    \alpha_s$ which means that a shorter path of delivery $\pi_{xz}'$
    exists that either forwards appropriate staleness notifications
    (A) or detects its possible blocking:
  \item
    $r_{y'}(\alpha_s) \wedge \delta_s \in D_{y'} \wedge \alpha_x =
    \last D_{y'}$ which detects a possible blocking of
    $\delta_x$. Without global knowledge, $y'$ assumes it belongs to
    the shortest and only path of delivery of $\alpha_x$
    ($\pi_{xy'} \subseteq \Pi_{xz}$) thus it does not further delegate
    the detection to another \process.
  \end{enumerate}
  \end{asparadesc}
  \vspace{-1.5em} %% ugly vspace but necessary for enumerate+qed
  %
  %% (i) If every \process (such as $a$) starting from the source
  %% delivers and forwards a removal notification $\delta_X$ when their
  %% last delivery is $\alpha_X$, then every such \process eventually
  %% delivers the removal notification $\delta_X$ except \processes (such
  %% as $b$) that delivered $\alpha_X$ from a \process that delivered a
  %% message from another partition $\alpha_Y$ since then.
  %
  %% These exceptions eventually receive and deliver $\alpha_Y$ since
  %% $\alpha_Y$ is better than $\alpha_X$ through this path, except if
  %% they already received and delivered the removal notification of
  %% $P_Y$ (such as $c$). These \processes may never receive hence
  %% deliver $\delta_X$, and may never receive hence deliver a better
  %% message than $\alpha_X$. They need an additional mechanism to
  %% eventually purge $\alpha_X$ that cannot rely on the eventual purging
  %% of $P_Y$ at \processes like $b$, to avoid deadlocks.
  %
  %% (ii) Assuming that every \process keeps an history of its past
  %% deliveries, \processes (such as $c$) can detect the inconsistency,
  %% since they receive from their parent an already deleted partition
  %% $P_Y$. This notification means that the parent discards any
  %% $\delta_Z$ with $\alpha_Z^z < \alpha_Y^y$, and most importantly, if
  %% $\delta_X$ exists, it discards it. To ensure the purge of stale
  %% messages, a detecting \process must assume the worst case that such
  %% $\delta_X$ exists, and send another kind of message, noted $\Delta$,
  %% that notify the possible removal of $P_X$.
  %
  %% (iii) A \process (such as $d$) whose last delivery is $\alpha_X$,
  %% but whose parent is neither inconsistent (like $b$) nor receiving
  %% $\delta_X$ (like $a$), eventually receives $\Delta$ from a detecting
  %% \process, either directly or transitively, for such a child \process
  %% ($d$) trusts the possible removal of $P_X$ by delivering and
  %% forwarding such $\Delta$. $\Delta$ suffers from identical blocking
  %% conditions (between $b$ and $c$) than $\delta$, leading to the same
  %% solutions of detection and propagation of $\Delta$. Eventually,
  %% every \process whose last delivery is $\alpha_X$ receives and
  %% delivers either a better $\alpha_Z$, or $\delta_X$, or $\Delta_X$.
\end{proof}
